---
title: "Methodology"
author: "Everet Rummel"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../../')

library(knitr)
```

```{r import, include=FALSE}
library(readr)

diversity_density = read_csv('./data/output/NJ_20132017_diversity_density.csv')

detach('package:readr', unload=TRUE)
```


# Background

*New Jersey Family* magazine created a list of ["New Jersey's Best Towns for Families"](https://www.njfamily.com/New-Jerseys-Best-Towns-for-Families/) 
(full list [here](https://www.njfamily.com/new-jerseys-best-towns-for-families-the-list-2019/). 
An [obvious question](https://twitter.com/hobokenurbanist/status/1114592402336100355) 
arose: How dense and racially diverse are the top towns in this list?

# Objectives

Using most recent, highest quality data available:

* Calculate a measure of racial diversity for all municipalities in New Jersey
* Calculate population density for all municipalities in New Jersey
* Calculate diversity/density rank and percentile for all municipalities

# Code

I did everything in `code/R/calculate.R`. See that script for all the details.
See below for a summary in English.

# Source Data

American Community Survey, 5-year Public Use Micro Sample, 2013-2017. The Census
Bureau provides estimates for county subdivisions, which basically equal 
municipalities, so I use those summary tables.

## Variables

### Race, Population

[Table B03002: HISPANIC OR LATINO ORIGIN BY RACE](https://factfinder.census.gov/bkmk/table/1.0/en/ACS/17_5YR/B03002/0400000US34.06000).

### Land area

[American FactFinder](https://factfinder.census.gov/bkmk/table/1.0/en/DEC/10_SF1/GCTPH1.ST16/0400000US34) 
only provides *rounded* land area. For more accurate land area data, I 
downloaded the [2017 TIGER/Line shapefile](https://www.census.gov/geographies/mapping-files/time-series/geo/tiger-line-file.2017.html) 
for New Jersey (see: [direct download link](https://www2.census.gov/geo/tiger/TIGER2017/COUSUB/tl_2017_34_cousub.zip)).

# Universe

Each record is a county subdivision, which I'm calling a "municipality" here.

I exclude municipalities with populations lower than the median, 8,244.

# Calculations

To measure diversity, I calculate the [Gini-Simpson Index](https://en.wikipedia.org/wiki/Diversity_index#Gini%E2%80%93Simpson_index).
If any economists are reading this, the Gini-Simpson Index is one minus the [Herfindahl–Hirschman Index](https://en.wikipedia.org/wiki/Herfindahl_index).
Hereafter, I'll refer to it as the "Diversity Index" in title-case.

The value of the Diversity Index for a given municipality can be interpreted as 
the probability that two randomly selected residents are of a different race.
For example: Two random residents in Jersey City have a ~75% chance of 
identifying as different races.

The Diversity Index is calculate as follows:

$D = 1 - \sum_{i = 1}^{N} p_i$

where $p_i$ is the percent of the total municipality population identifying as
racial group $i$, and $N$ is the total number of racial groups. 

There are 7 racial groups in the Census: White, Black or African American,
American Indian and Alaska Native, Asian, Native Hawaiian and Other Pacific
Islander, Two or more races, or Some other race. If you glance at the 
[HISPANIC OR LATINO ORIGIN BY RACE](https://factfinder.census.gov/bkmk/table/1.0/en/ACS/17_5YR/B03002/0400000US34.06000) table, you'll notice many people identifying with Hispanic ethnicity
identify as "Some other race". 

## Special Note About Hispanic Individuals

Because "Hispanic" is an ethnicity, not a racial group (according to the Census
Bureau's admittedly imperfect definitions), 
**I do not use the Hispanic variable to calculate the Diversity Index.** In 
doing so, I'm not leaving out Hispanic people. Most of the people identifying as
"some other race" are Hispanic. All other Hispanic people identify as white, 
black, asian, etc, and so they're included in the count of the race with which
they identify.

# Validation

My results can be compared to the Diversity Index calculated for every 
municipality in New Jersey by NJ.com's [Disha Raychaudhuri](https://twitter.com/Disha_RC) 
in [this article](https://www.nj.com/data/2019/02/the-25-most-racially-diverse-towns-in-nj-ranked.html). 
My rankings match theirs almost exactly. My calculations are off by about a
hundredth of a percentage point for each municipality. They claim to have 
followed [this methodology](https://www.usatoday.com/story/news/nation/2014/10/21/diversity-index-data-how-we-did-report/17432103/) 
for calculating the Diversity Index from ACS data, but I matched their results 
by *not* following that methodology.

Note: While Raychaudhuri excludes municipalities with populations lower than 
10,000, I exclude those with populations lower than the median, 8,244.

# My Environment

Here are the R packages directly loaded in this analysis:

```{r}
library(dplyr)
library(readr)
library(scales)
library(sf)
```

Here is my session info:

```{r}
sessionInfo()
```

# Results

The full cleaned dataset output by `calculate.R` is in `data/output/NJ_20132017_diversity_density.csv`. The following results come
from this dataset, unless stated otherwise.

The 10 most diverse municipalities in New Jersey:

```{r, echo=FALSE, message=FALSE}
diversity_density %>%
  head(n = 10) %>%
  knitr::kable()
```

The 10 least diverse municipalities in New Jersey:

```{r, echo=FALSE, message=FALSE}
diversity_density %>%
  tail(n = 10) %>%
  knitr::kable()
```

The 10 densest municipalities:

```{r, echo=FALSE, message=FALSE}}
diversity_density %>%
  arrange(-`Population density`) %>%
  head(n = 10) %>%
  knitr::kable()
```

The 10 sparsest:

```{r, echo=FALSE, message=FALSE}}
diversity_density %>%
  arrange(-`Population density`) %>%
  tail(n = 10) %>%
  knitr::kable()
```
