---
output: github_document
title: "Methodology"
author: "Everet Rummel"
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Source data notes

The ACS 2013-17 sample covers years *2012* to 2017.

## Universe

The ACS has county subdivisions, which I usually call "municipalities" here. I
exclude "County subdivisions not defined" records.

In all datasets with rankings, I exclude municipalities with populations lower
than 1,000. In the past, I used the median town population as the cutoff, but
that seemed a bit high, and doing so excluded some interesting findings from the
results.

## Calculations

To measure diversity, I calculate the [Gini-Simpson Index](https://en.wikipedia.org/wiki/Diversity_index#Gini%E2%80%93Simpson_index).
If any economists are reading this, the Gini-Simpson Index is one minus the
[Herfindahl–Hirschman Index](https://en.wikipedia.org/wiki/Herfindahl_index).
Hereafter, I'll refer to it as the "Diversity Index" in title-case.

The value of the Diversity Index for a given municipality can be interpreted as 
the probability that two randomly selected residents are of a different race.
For example: Two random residents in Jersey City have a ~75% chance of 
identifying as different races.

The Diversity Index is calculated as follows:

$D = 1 - \sum_{i = 1}^{N} p_i$

where $p_i$ is the percent of the total municipality population identifying as
racial group $i$, and $N$ is the total number of racial groups. 

There are 7 racial groups in the ACS: White, Black or African American, American
Indian and Alaska Native, Asian, Native Hawaiian and Other Pacific Islander, Two
or more races, and Some other race. Because "Hispanic" is not a racial group
(according to the Census Bureau's admittedly imperfect definitions),
**I do not use the Hispanic grouping to calculate the Diversity Index.** In 
doing so, I'm not leaving out Hispanic people. If you glance at the 
["HISPANIC OR LATINO ORIGIN BY RACE"](https://factfinder.census.gov/bkmk/table/1.0/en/ACS/17_5YR/B03002)
summary table, you'll notice most people identifying as "some other race" are
Hispanic/Latinx in origin. All other Hispanic/Latinx people identify as White,
Black, Asian, etc, and so they're included in the count of the race with which
they identify.

## Validation

My racial diversity results can be compared to those calculated by NJ.com's 
[Disha Raychaudhuri](https://twitter.com/Disha_RC)
in
[this article](https://www.nj.com/data/2019/02/the-25-most-racially-diverse-towns-in-nj-ranked.html).
My rankings match theirs almost exactly. My calculations are off by about a
hundredth of a percentage point for each municipality. They claim to have 
followed
[this methodology](https://www.usatoday.com/story/news/nation/2014/10/21/diversity-index-data-how-we-did-report/17432103/) 
for calculating the Diversity Index from ACS data, but I matched their results 
by *not* following that methodology. See [my code](code/diversity_race.R) for
details.

My economic diversity results can be compared to those calculated by NJ.com's
[Disha Raychaudhuri](https://twitter.com/Disha_RC)
in
[this article](https://www.nj.com/data/2019/04/nj-towns-are-increasingly-becoming-rich-or-poor-is-the-middle-class-disappearing.html).
Once again, I'm able to almost exactly match their results.

Note: While Raychaudhuri excludes municipalities with populations lower than 
10,000, I exclude those with populations lower than 1,000.

## My environment

Here is my session info after running all lines of `__plan__.R`:

```{r}
sessionInfo()
```
